Supplementary Material for Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-05-14 supp_mat_complexity_selection.Rmd
update supp mat

@@

GAM estimates of selection gradients differ from GAMM

```{r COMPARE GAM WITH GAMM}

## CONTROL PLANTS ----

## Gall Height
gam_red.control_size <- gam(gall_survival ~
                             s(c.Gall_Height_mm) + 
                             c.gall_individuals +
                             c.Density_per_100_shoots,
                           data = control_df, family = binomial(link = "logit"))
summary(gam_red.control)

gam.plot(gam_red.control)
gam.plot(gamm_red.control$gam)

## Gall individuals
control_df_gall.level <- control_df %>%
  group_by(Treatment.focus, Plant_Position, Gall_Number) %>%
  summarise(gall_survival_mean = mean(gall_survival),
            gall_individuals = mean(gall_individuals),                 # note that there is no variance, it is the actual value
            Gall_Height_mm_mean = mean(Gall_Height_mm),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>% # remember that this is pseudo-replicated at the gall-level
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals = gall_individuals - mean(gall_individuals),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.control_indiv <- gam(gall_survival_mean ~
                              c.Gall_Height_mm_mean + 
                              s(c.gall_individuals, k=9) +
                              c.Density_per_100_shoots,
                             data = control_df_gall.level, family = binomial(link = "logit")) # we didn't include weights because this would be correlated with the number of individuals per gall
summary(gam_red.control_indiv)

gam.plot(gam_red.control_indiv)
gam.plot(gamm_red.control$gam)

## Gall density
control_df_plant.level <- control_df_gall.level %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise(gall_survival_mean = mean(gall_survival_mean),
            gall_sample_size = n(),
            gall_individuals_mean = mean(gall_individuals),                 
            Gall_Height_mm_mean = mean(Gall_Height_mm_mean),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%                # note that there is no variance, it is the actual value
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.control_density <- gam(gall_survival_mean ~
                                c.Gall_Height_mm_mean + 
                                c.gall_individuals_mean +
                                s(c.Density_per_100_shoots, k=3),   # do I adjust k to be consistent with mixed-effects model?
                             data = control_df_plant.level, family = binomial(link = "logit"), weights = gall_sample_size) # weighting appears to alter the results...
summary(gam_red.control_density)

gam.plot(gam_red.control_density)
gam.plot(gamm_red.control$gam)

## TREATMENT PLANTS ----

## Gall Height
gam_red.treatment_size <- gam(gall_survival ~
                         s(c.Gall_Height_mm) + 
                         c.gall_individuals +
                         c.Density_per_100_shoots,
                       data = treatment_df, family = binomial(link = "logit"))
summary(gam_red.treatment)

gam.plot(gam_red.treatment)
gam.plot(gamm_red.treatment$gam)

## Gall individuals
treatment_df_gall.level <- treatment_df %>%
  group_by(Treatment.focus, Plant_Position, Gall_Number) %>%
  summarise(gall_survival_mean = mean(gall_survival),
            gall_individuals = mean(gall_individuals),                 # note that there is no variance, it is the actual value
            Gall_Height_mm_mean = mean(Gall_Height_mm),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>% # remember that this is pseudo-replicated at the gall-level
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals = gall_individuals - mean(gall_individuals),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.treatment_indiv <- gam(gall_survival_mean ~
                              c.Gall_Height_mm_mean + 
                              s(c.gall_individuals, k=9) +
                              c.Density_per_100_shoots,
                             data = treatment_df_gall.level, family = binomial(link = "logit"))
summary(gam_red.treatment_indiv)

gam.plot(gam_red.treatment_indiv)
gam.plot(gamm_red.treatment$gam)

## Gall density
treatment_df_plant.level <- treatment_df_gall.level %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise(gall_survival_mean = mean(gall_survival_mean),
            gall_sample_size = n(),
            gall_individuals_mean = mean(gall_individuals),                 
            Gall_Height_mm_mean = mean(Gall_Height_mm_mean),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%                # note that there is no variance, it is the actual value
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.treatment_density <- gam(gall_survival_mean ~
                                c.Gall_Height_mm_mean + 
                                c.gall_individuals_mean +
                                s(c.Density_per_100_shoots),
                             data = treatment_df_plant.level, family = binomial(link = "logit"), weights = gall_sample_size) # weighting appears to alter the results...
summary(gam_red.treatment_density)

gam.plot(gam_red.treatment_density)
gam.plot(gamm_red.treatment$gam)

#### CALCULATE SELECTION GRADIENTS FROM GAM ----

## GALL SIZE ----
sg.gam.size_control <- gradient.calc(mod = gam_red.control, 
                                 phenotype = "c.Gall_Height_mm", 
                                 covariates = c("c.gall_individuals","c.Density_per_100_shoots"))
sg.gam.size_treatment <- gradient.calc(mod = gam_red.treatment, 
                                   phenotype = "c.Gall_Height_mm", 
                                   covariates = c("c.gall_individuals","c.Density_per_100_shoots"))

## GALL INDIVIDUALS ----
sg.gam.individuals_control <- gradient.calc(mod = gam_red.control, 
                                        phenotype = "c.gall_individuals", 
                                        covariates = c("c.Gall_Height_mm","c.Density_per_100_shoots"))
sg.gam.individuals_treatment <- gradient.calc(mod = gam_red.treatment, 
                                          phenotype = "c.gall_individuals", 
                                          covariates = c("c.Gall_Height_mm","c.Density_per_100_shoots"))

## GALL DENSITY ----
sg.gam.density_control <- gradient.calc(mod = gam_red.control, 
                                    phenotype = "c.Density_per_100_shoots", 
                                    covariates = c("c.gall_individuals","c.Gall_Height_mm"))
sg.gam.density_treatment <- gradient.calc(mod = gam_red.treatment, 
                                      phenotype = "c.Density_per_100_shoots", 
                                      covariates = c("c.gall_individuals","c.Gall_Height_mm"))

## PRINT TABLE ----
rbind(
  mutate(sg.gam.size_control$ests, Food_Web_Treatment = "Complex"),
  mutate(sg.gam.size_treatment$ests, Food_Web_Treatment = "Simple"),
  mutate(sg.gam.individuals_control$ests, Food_Web_Treatment = "Complex"),
  mutate(sg.gam.individuals_treatment$ests, Food_Web_Treatment = "Simple"),
  mutate(sg.gam.density_control$ests, Food_Web_Treatment = "Complex"),
  mutate(sg.gam.density_treatment$ests, Food_Web_Treatment = "Simple")
)

round(rbind(
  sg.gam.size_control$ests,
  sg.gam.size_treatment$ests,
  sg.gam.individuals_control$ests,
  sg.gam.individuals_treatment$ests,
  sg.gam.density_control$ests,
  sg.gam.density_treatment$ests
),3)

## BOOTSTRAPPED INTERVALS ----
gam.bootstrap_df <- bind_rows(
  data.frame(Bootstrap_Estimates = sg.gam.size_control$boot[ ,1], Phenotype = "c.Gall_Height_mm", Selection_Type = "Beta", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.size_control$boot[ ,2], Phenotype = "c.Gall_Height_mm", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.size_treatment$boot[ ,1], Phenotype = "c.Gall_Height_mm", Selection_Type = "Beta", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.size_treatment$boot[ ,2], Phenotype = "c.Gall_Height_mm", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.individuals_control$boot[ ,1], Phenotype = "c.gall_individuals", Selection_Type = "Beta", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.individuals_control$boot[ ,2], Phenotype = "c.gall_individuals", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.individuals_treatment$boot[ ,1], Phenotype = "c.gall_individuals", Selection_Type = "Beta", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.individuals_treatment$boot[ ,2], Phenotype = "c.gall_individuals", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.density_control$boot[ ,1], Phenotype = "c.Density_per_100_shoots", Selection_Type = "Beta", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.density_control$boot[ ,2], Phenotype = "c.Density_per_100_shoots", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.density_treatment$boot[ ,1], Phenotype = "c.Density_per_100_shoots", Selection_Type = "Beta", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.density_treatment$boot[ ,2], Phenotype = "c.Density_per_100_shoots", Selection_Type = "Gamma", Food_Web_Treatment = "Simple")
)

gam.bootstrap_df %>%
  ggplot(., aes(x=Phenotype, y=Bootstrap_Estimates, fill=Food_Web_Treatment)) +
  geom_boxplot() +
  geom_hline(yintercept=0, linetype="dotted") +
  facet_wrap(~Selection_Type) +
  coord_flip()

>>>>

```{r}
sg.size.x.individuals_control <- gradient.calc(mod = gamm_red.control$gam, 
                                               phenotype = c("c.Gall_Height_mm", "c.gall_individuals"),
                                               covariates = c("c.Density_per_100_shoots"))
sg.size.x.individuals_treatment <- gradient.calc(mod = gamm_red.treatment$gam, 
                                                 phenotype = c("c.Gall_Height_mm", "c.gall_individuals"),
                                                 covariates = c("c.Density_per_100_shoots"))
sg.size.x.density_control <- gradient.calc(mod = gamm_red.control$gam, 
                                               phenotype = c("c.Gall_Height_mm", "c.Density_per_100_shoots"),
                                               covariates = c("c.gall_individuals"))
sg.size.x.density_treatment <- gradient.calc(mod = gamm_red.treatment$gam, 
                                                 phenotype = c("c.Gall_Height_mm","c.Density_per_100_shoots"),
                                                 covariates = c("c.gall_individuals"))
sg.density.x.individuals_control <- gradient.calc(mod = gamm_red.control$gam, 
                                                  phenotype = c("c.Density_per_100_shoots", "c.gall_individuals"),
                                                  covariates = c("c.Gall_Height_mm"))
sg.density.x.individuals_treatment <- gradient.calc(mod = gamm_red.treatment$gam, 
                                                    phenotype = c("c.Density_per_100_shoots", "c.gall_individuals"),
                                                    covariates = c("c.Gall_Height_mm"))

# note that while the means are always the same for each gradient calculation, the SE and p-values can be slightly different. I always reported the highest SE and highest P-value to be conservative
bind_rows(data.frame(round(sg.size.x.individuals_control$ests,3), 
                     gradients = rownames(sg.size.x.individuals_control$ests),
                     treatment = "Complex"),
          data.frame(round(sg.size.x.individuals_treatment$ests,3), 
                     gradients = rownames(sg.size.x.individuals_treatment$ests),
                     treatment = "Simple"),
          data.frame(round(sg.size.x.density_control$ests,3), 
                     gradients = rownames(sg.size.x.density_control$ests),
                     treatment = "Complex"),
          data.frame(round(sg.size.x.density_treatment$ests,3), 
                     gradients = rownames(sg.size.x.density_treatment$ests),
                     treatment = "Simple"),
          data.frame(round(sg.density.x.individuals_control$ests,3), 
                     gradients = rownames(sg.density.x.individuals_control$ests),
                     treatment = "Complex"),
          data.frame(round(sg.density.x.individuals_treatment$ests,3), 
                     gradients = rownames(sg.density.x.individuals_treatment$ests),
                     treatment = "Simple")) %>%
  distinct %>%
  arrange(gradients)

@@

Now we plot the fitness landscapes associated with different combinations of gall phenotypes. We restrict these plots to 1 SD above and below the mean trait value, as this contains the majority of phenotypes under selection. Therefore, this fitness landscape integrates information from the fitness function with information about distributions of phenotype, and has been suggested to be a more informative way to visualize the magnitude of natural selection.
```{r Fitness Landscapes}

#### ESTIMATE FITNESS LANDSCAPES ----

## GALL SIZE
fl.size_control <- insect_FL(mod = gamm_red.control$gam, 
                                 phenotype = "c.Gall_Height_mm", 
                                 covariates = c("c.gall_individuals","c.Density_per_100_shoots"))
fl.size_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                   phenotype = "c.Gall_Height_mm", 
                                   covariates = c("c.gall_individuals","c.Density_per_100_shoots"))

## GALL INDIVIDUALS
fl.individuals_control <- insect_FL(mod = gamm_red.control$gam, 
                                        phenotype = "c.gall_individuals", 
                                        covariates = c("c.Gall_Height_mm","c.Density_per_100_shoots"))
fl.individuals_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                          phenotype = "c.gall_individuals", 
                                          covariates = c("c.Gall_Height_mm","c.Density_per_100_shoots"))

## GALL DENSITY 
fl.density_control <- insect_FL(mod = gamm_red.control$gam, 
                                    phenotype = "c.Density_per_100_shoots", 
                                    covariates = c("c.gall_individuals","c.Gall_Height_mm"))
fl.density_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                      phenotype = "c.Density_per_100_shoots", 
                                      covariates = c("c.gall_individuals","c.Gall_Height_mm"))

## GALL INDIVIDUALS X DENSITY
fl.individuals.x.density_control <- insect_FL(mod = gamm_red.control$gam, 
                                              phenotype = c("c.Density_per_100_shoots","c.gall_individuals"), 
                                              covariates = c("c.Gall_Height_mm"))
fl.individuals.x.density_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                                phenotype = c("c.Density_per_100_shoots","c.gall_individuals"), 
                                                covariates = c("c.Gall_Height_mm"))

## GALL INDIVIDUALS x SIZE
fl.individuals.x.size_control <- insect_FL(mod = gamm_red.control$gam, 
                                            phenotype = c("c.Gall_Height_mm","c.gall_individuals"), 
                                            covariates = c("c.Density_per_100_shoots"))
fl.individuals.x.size_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                             phenotype = c("c.Gall_Height_mm","c.gall_individuals"), 
                                             covariates = c("c.Density_per_100_shoots"))

## GALL DENSITY X SIZE
fl.size.x.density_control <- insect_FL(mod = gamm_red.control$gam, 
                                       phenotype = c("c.Density_per_100_shoots","c.Gall_Height_mm"), 
                                       covariates = c("c.gall_individuals"))
fl.size.x.density_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                         phenotype = c("c.Density_per_100_shoots","c.Gall_Height_mm"), 
                                         covariates = c("c.gall_individuals"))

#### PLOT FITNESS LANDSCAPES ----

# Order: orange, light blue, green, yellow, dark blue, red, pink
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000", "#999999")
complex_color <- cbPalette[6]
simple_color <- cbPalette[5]
treatmeant_colors <- c(complex_color, simple_color)

## 1-DIMENSIONAL
uni_size <- bind_rows(mutate(get_insect_FL.df(fl.size_control), Treatment = "Complex"),
          mutate(get_insect_FL.df(fl.size_treatment), Treatment = "Simple")) %>%
  ggplot(.) +
  geom_line(aes(x = Var1, y = mean_fitness, color=Treatment)) + 
  geom_ribbon(aes(x = Var1, ymin = lower_2.5, ymax = upper_97.5, fill=Treatment), alpha = 0.25) +
  ylab("Mean fitness") +
  xlab("Mean gall size") +
  scale_color_manual(values = treatmeant_colors) +
  scale_fill_manual(values = treatmeant_colors) +
  scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1), limits = c(0,1))

uni_clutch <- bind_rows(mutate(get_insect_FL.df(fl.individuals_control), Treatment = "Complex"),
          mutate(get_insect_FL.df(fl.individuals_treatment), Treatment = "Simple")) %>%
  ggplot(.) +
  geom_line(aes(x = Var1, y = mean_fitness, color=Treatment)) + 
  geom_ribbon(aes(x = Var1, ymin = lower_2.5, ymax = upper_97.5, fill=Treatment), alpha = 0.25) +
  ylab("Mean fitness") +
  xlab("Mean clutch size") +
  scale_color_manual(values = treatmeant_colors) +
  scale_fill_manual(values = treatmeant_colors) +
  scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1), limits = c(0,1))

uni_pref <- bind_rows(mutate(get_insect_FL.df(fl.density_control), Treatment = "Complex"),
          mutate(get_insect_FL.df(fl.density_treatment), Treatment = "Simple")) %>%
  ggplot(.) +
  geom_line(aes(x = Var1, y = mean_fitness, color=Treatment)) + 
  geom_ribbon(aes(x = Var1, ymin = lower_2.5, ymax = upper_97.5, fill=Treatment), alpha = 0.25) +
  ylab("Mean fitness") +
  xlab("Mean female preference") +
  scale_color_manual(values = treatmeant_colors) +
  scale_fill_manual(values = treatmeant_colors) +
  scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1), limits = c(0,1))

uni_legend <- get_legend(uni_size)

uni_plots <- plot_grid(uni_size + theme(legend.position = "none"), 
                       uni_clutch + theme(legend.position = "none", 
                                          axis.title.y = element_blank(), axis.text.y = element_blank()),
                       uni_pref + theme(legend.position = "none", 
                                        axis.title.y = element_blank(), axis.text.y = element_blank()),
                       labels = "AUTO", nrow = 1, align='hv')
plot_grid(uni_plots, uni_legend, ncol=2, rel_widths = c(3,0.3))

## 2-DIMENSIONAL

bind_rows(mutate(get_insect_FL.df(fl.individuals.x.density_treatment), Treatment="Simple"),
          mutate(get_insect_FL.df(fl.individuals.x.density_control), Treatment="Complex")) %>%
  ggplot(., aes(x=Var1, y=Var2, fill=mean_fitness)) +
  geom_raster() + 
  facet_wrap(~Treatment) +
  scale_fill_viridis() +
  xlab("Female preference") +
  ylab("Clutch size")

bind_rows(mutate(get_insect_FL.df(fl.individuals.x.size_treatment), Treatment="Simple"),
          mutate(get_insect_FL.df(fl.individuals.x.size_control), Treatment="Complex")) %>%
  ggplot(., aes(x=Var1, y=Var2, fill=mean_fitness)) +
  geom_raster() + 
  facet_wrap(~Treatment) +
  scale_fill_viridis() +
  xlab("Clutch size") +
  ylab("Gall size")

bind_rows(mutate(get_insect_FL.df(fl.size.x.density_treatment), Treatment="Simple"),
          mutate(get_insect_FL.df(fl.size.x.density_control), Treatment="Complex")) %>%
  ggplot(., aes(x=Var1, y=Var2, fill=mean_fitness)) +
  geom_raster() + 
  facet_wrap(~Treatment) +
  scale_fill_viridis() +
  xlab("Gall size") +
  ylab("Female preference")
```
